---
title: "R - Climate Data the world bank"
author: "Frank Liang"
date: "27/07/2021"
output: html_document
---

```{r}
library(tidyverse)
library(ggplot2)

#install.packages(c("httr", "jsonlite"))
library(httr)
library(jsonlite)
```

# World Bank API
Note a lot of the data are forecast based on models
details regard api visit: https://datahelpdesk.worldbank.org/knowledgebase/articles/902061-climate-data-api

The Climate Data API uses REST-based requests, in which the general form looks like this:

`http://climatedataapi.worldbank.org/climateweb/rest/v1/country/type/var/start/end/ISO3`

`mavg` for Monthly average 
`annualavg` for Annual average 
`manom` average monthly change
`annualanom` average annual change

Two variable you can study about flood by `precipitation` or by `basinID`. for latter just change the country code to a number from 1 to 468. [Click here](http://data.worldbank.org/sites/default/files/climate_data_api_basins.pdf) for a complete basin map.

```{r}
link = "http://climatedataapi.worldbank.org/climateweb/rest/v1/country/mavg/pr/2020/2039/CHN.csv"

data <- read.csv(link)
View(data)
data[,c(6:17)]
```
In this data frame the 15 GCM are types of circulation models used to estimate... 


# R- NOAA data
thanks to [this data](http://spatialecology.weebly.com/r-code--data/34)

```{r}
YourTokenHere = "xOhVdMrdWTsFSZXCrRZJPXplNlPbgUWr"
```

```{r}
# Enter the Token that was sent to your email below.
options(noaakey = "xOhVdMrdWTsFSZXCrRZJPXplNlPbgUWr")
require(devtools)  #Make sure this package is installed
devtools::find_rtools()  #Make sure Rtools can be found

devtools::install_github("ropensci/rnoaa")  # Installing the rnoaa() package from the GitHub Page.
```

```{r}
library(rnoaa)  #Now you can load all the required libraries.
library(ncdf)
#install.packages("ncdf")
library(plyr)
library(devtools)

#install.packages("fields")
library(fields)
#install.packages("GEOmap")
library(GEOmap)
```

```{r}
MI_Monthly <- ncdc(datasetid = "GHCNDMS", locationid = "FIPS:26065", stationid = "GHCND:USW00014836", 
    datatypeid = "TPCP", startdate = "2000-01-01", enddate = "2010-01-01", limit = 1000, token = YourTokenHere)

ncdc_plot(MI_Monthly, breaks = "12 months", dateformat = "%Y")
```

```{r}
MI_Annual <- ncdc(datasetid = "ANNUAL", locationid = "FIPS:26065", datatypeid = "DSNW", 
    startdate = "2000-01-01", enddate = "2010-01-01", limit = 1000, token = YourTokenHere)
ncdc_plot(MI_Annual, breaks = "1 year", dateformat = "%Y")
```

```{r}
# list of all stations:
ghcnd_stations(lattitude = 53.4808 , longitude = -2.2426, first_year = 2015, last_year = 2020) -> stations

#make df
lat_lon_df <- data.frame(id = c("sydney", "brisbane"),
                         latitude = c(-33.8675, -27.4710),
                         longitude = c(151.2070,153.0234))
meteo_nearby_stations(
  lat_lon_df = lat_lon_df,
  station_data = stations, 
  radius = 50,
  limit = 10
) -> nearStation
```

```{r}
library(leaflet)
```

```{r}
nearStation$sydney %>%
  View()
```

```{r}
leaflet(nearStation$sydney) %>%
  setView(lat = -33.8675, lng = 151.2070, zoom = 12) %>% 
  addProviderTiles(providers$CartoDB.Positron) %>%
  addMarkers(popup = ~id, label = ~name)

```

