---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r message=FALSE, include=FALSE}
library(rnoaa)  #Now you can load all the required libraries.
#library(ncdf)
#install.packages("ncdf")
library(plyr)
library(dplyr)
library(readr)
library(ggplot2)
library(plotly)

library(tidyverse)
```

# Risks

## Possion Process & Linear Model
```{r}
library(dplyr)
library(readr)
df <- read_csv("raw/prcp.csv")

df %>%
  mutate(year = format(date, "%Y")) %>%
  dplyr::select(date, year, prcp) %>%
  group_by(year) %>%
  filter(prcp != 0) %>%
  summarise(rain_days = length(prcp)) -> rainDays

df %>%
  mutate(year = format(date, "%Y")) %>%
  dplyr::select(date, year, prcp) %>%
  group_by(year) %>%
  summarise(year_days = length(year)) -> yearDays
```


```{r}
# filter out year that contains irregular day of years
yearDays %>%
  filter(year_days < 365)
```


```{r}
rainDays %>% left_join(yearDays, by = "year") %>%
  mutate(rate = rain_days/year_days) %>%
  with(plot(x = year, y = rate))
```

```{r}
rainDays %>% left_join(yearDays, by = "year") %>%
  mutate(rate = rain_days/year_days) %>%
  pull(rate) %>%
  density() %>%
  plot()
```


```{r}
rainDays %>% left_join(yearDays, by = "year") %>%
  mutate(rate = rain_days/year_days) %>%
  filter(year <= 2015) %>%
  mutate(year = as.numeric(year)) -> data
length(data$year)
```

```{r}
sd(data$rate)
mean(data$rate)

lm(data = data, rate ~ year) -> mod.lm
summary(mod.lm)
residuals(mod.lm) %>% density() %>% plot
```
mean of rain rate = 0.2217114
sd of rain rate = 0.03215555
model sd = 0.03056
F-statistic: 6.788 on 1 and 51 DF,  p-value: 0.012

In this statistically model you assume the change of rain is normally distributed. But in this case residual is slightly tiled towards left... 

So... 85% confident that the frequency of raining days are slightly decreasing

## Expotential 
Signs that percepitation is exponentially distributed. 

```{r}
library(dplyr)
df %>%
  mutate(year = format(date, "%Y")) %>%
  dplyr::select(date, year, prcp) %>%
  group_by(year) %>%
  filter(prcp != 0) %>%
  filter(year <= 2012) %>%
  summarise(mean = mean(prcp), sd = sd(prcp), median = median(prcp)) -> x

with(x, plot(x = mean, y = sd, sub = "\nStandard deviation is linear to that mean value each year. \nSuggesting exponential distribution"))
with(x, plot(x = mean, y = median))
with(x, plot(x = year, y = mean - median))
with(x, plot(x = mean, y = sd/mean))

lm(data = x, sd ~ mean) %>% summary()
```
This should be enough to prove the potential distribution because mean and is correlated??

```{r}
df %>%
  mutate(year = format(date, "%Y")) %>%
  dplyr::select(date, year, prcp) %>%
  group_by(year) %>%
  filter(prcp != 0) %>%
  filter(year <= 2012) %>%
  mutate(log_prc = log(prcp)) %>%
  summarise(mean = mean(log_prc), sd = sd(log_prc)) -> log_x

with(log_x, plot(x = mean, y = sd))

df %>%
  mutate(year = format(date, "%Y")) %>%
  dplyr::select(date, year, prcp) %>%
  group_by(year) %>%
  filter(prcp != 0) %>%
  filter(year <= 2012) %>%
  mutate(log_prc = 5.041121*log(prcp)) %>%
  summarise(mean = mean(log_prc), sd = sd(log_prc)) -> logk_x
with(logk_x, plot(x = mean, y = sd))

```


```{r}
df %>%
  mutate(year = format(date, "%Y")) %>%
  dplyr::select(date, year, prcp) %>%
  group_by(year) %>%
  filter(prcp != 0) %>%
  filter(year <= 2012) %>%
  pull(prcp) %>%
  hist(breaks = 1000, freq = TRUE) -> hist
names(hist)
hist$density %>% length()
hist$mids %>% length()

x <- hist$mids
y <- hist$density

data.frame(x, y) -> seed
seed[sample(nrow(seed), 750),]

library(drc)
library(nlme)

update.packages("drc")
seed %>%
  filter( x <= 400) -> tailess

mod.exd <- drm(y ~ x, fct = EXD.2(), data = seed)
mod <- drm(y ~ x, fct = AR.2(), data = seed)

```
plot hist
```{r}
with(hist, plot(mids, density, pch = ".", xlim = c(0, 200)))
sum(hist$density)
```
plot model
```{r}
plot(x = seed$x, y = fitted(mod.exd), type = "l", col = "blue", xlim = c(0,300))
lines(x = seed$x, y = tailess$y, type = "p", col = alpha("black", 0.5))
```
```{r}
model <- drm(y ~ log(x), fct = EXD.2(), data = seed)
```

```{r}
plot(x = seed$x, y = fitted(model), type = "l", col = "blue", xlim = c(0, 300))
lines(x = seed$x, y = seed$y, type = "p", col = alpha("black", 0.5))
```

So you it proves log(x) is exponentially distributed? 
```{r}
summary(model)
```

f(x) = c + (d-c)(\exp(-x/e))

0.7828405 = 1/e = Ighag = 10* 0.07740550
 I think it is something to do with the scale x needs to be multiplied by 10 
 
 or log x needs to be multiplied by 10 
 
so we derive this function perhaps?? 10 * log (x) ~ exponential() and then give parameter a value 0.07740550.. 

Hist density is a function??

```{r}
x = log()
d = 0.07740550
e = 1.27739939

k <- 1/e/d  #10.1135
log(1000)*k

1 - pexp(log(1000)*k, d)
```
compare this result to from counting raindays 
```{r}
df %>%
  mutate(year = format(date, "%Y")) %>%
  dplyr::select(date, year, prcp) %>%
  group_by(year) %>%
  filter(prcp != 0) %>%
  pull(prcp) %>%
  length()

%>%
  filter(year <= 2012) %>%
  filter(prcp >= 1000) 

%>%
  pull(prcp)  %>%
  length()


13/5025

x <- hist$mids
y <- hist$count

data.frame(x, y) -> seed
seed[seed$x >= 1000,]$y

sum(hist$counts)
length(df$prcp)
```
0.002587065 0.25% percent... 

compares to probability density estimate: 0.004482068

0.4% probability that in any rain day a rain would exceed 10cm 


```{r}

```



# Prediction
```{r}
#library(forecast)
g <- df %>%
  ggplot(aes(x = date, y = prcp)) + 
  geom_line(color = "blue") + 
  theme(panel.border = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_rect(fill = "#fafafa"),
          plot.background = element_rect(fill = "white")
        )
ggplotly(g)
```




















